Energy and exergy analysis of a cruise ship

### 2017-07-03 Manuscript/Manuscript.tex
Update on the Python files for data analysis
Various extensions to Python files

Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse. The assumption was validated using for the June-December period, for which onboard measurements were available. This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
% We did the calculation, resampling the data from SHMI from a 1(3)h to 15min frequency using linear interpolation.

@@

\subsubsection*{Consumers-side}
Following, the assumptions that we have made related to the bottom-up estimations of ship heat demand:
\begin{itemize}
	\item The maximum demand for \textbf{spacial heating} is connected to a design minimum outer air temperature of -5$^o$C.
	\item The minimum demand for \textbf{spacial heating} is 0 for an external temperature of 20$^o$C (NOTE: Maybe 15 is more reasonable?)
	\item \textbf{Spacial heating} demand is linearly dependent on the external air temperature
	\item In the HVAC unit, the "\textbf{Preheater}" is only used when heating (i.e., in winter) while the "\textbf{Reheater}" is only used when cooling (i.e., in summer)
	\item Every passenger contributes with a total of 150 W of free heat during the day (8-23, ref. to "walking, seated" activity from \cite{Wang2000}) and 100 W during night (23-8, ref. to "seated at rest" activity from \cite{Wang2000}). For the crew, as it is assumed that they perform a more intense activity than the passengers, the daily contribution is considered equal to 160 (ref. to "moderate work" activity from \cite{Wang2000}) 
	\item The heat required for \textbf{tank heating} is linearly dependent on the external sea water temperature. The maximum heat is required for an external seawater temperature of 0$^o$C
	\item The heat required from the \textbf{galley} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. The hour-dependent factor is 0 at night (21-7), 0.1 during non-eating times (10-11 and 14-18) and 1 during eating times
	\end{itemize}
	\item The heat required from the \textbf{hot water} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. The hour-dependent factor is assumed based on the estimations for a land-based hotel reported in \cite{Chung2015}.
	\end{itemize}   
\end{itemize}

@@

>>>>

Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse. The assumption was validated based on  June-December period, for which onboard measurements were available. This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work.
% We did the calculation, resampling the data from SHMI from a 1(3)h to 15min frequency using linear interpolation.

@@

\subsubsection*{Consumers-side}
Following, the assumptions that we have made related to the bottom-up estimations of ship heat demand:
\begin{itemize}
	\item In the HVAC unit, the "\textbf{Preheater}" is only used when heating (i.e., in winter) while the "\textbf{Reheater}" is only used when cooling (i.e., in summer)
	\item \textbf{Spacial heating} demand during \textbf{winter} is linearly dependent on the external air temperature, from a minimum of 0 for an external temperature of 20$^o$C to a maximum of 3500 kW (design power of the preheater)  for an outer air temperature of -20$^o$C.
	\item \textbf{Spacial heating} demand during \textbf{summer} is related to the need for dehumidification, and harder to estimate based on available measurements. As a first assumption, it is estimated to be equal to the cooling demand (estimated as detailed in Sec. ref{sec:met:electric})
	\item Every passenger contributes with a total of 150 W of free heat during the day (8-23, ref. to "walking, seated" activity from \cite{Wang2000}) and 100 W during night (23-8, ref. to "seated at rest" activity from \cite{Wang2000}). For the crew, as it is assumed that they perform a more intense activity than the passengers, the daily contribution is considered equal to 160 (ref. to "moderate work" activity from \cite{Wang2000}) 
	\item The heat required for \textbf{tank heating} is linearly dependent on the external sea water temperature. The maximum heat is required for an external seawater temperature of 0$^o$C
	\item The heat required from the \textbf{galley} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. See Figure \ref{fig:hwhvstime}
	\end{itemize}
	\item The heat required from the \textbf{hot water} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. The hour-dependent factor is assumed based on the estimations for a land-based hotel reported in \cite{Chung2015}.
	\end{itemize}   
\end{itemize}

@@

\subsection{Estimation of the electric power demand} \label{sec:met:electric}

The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
\begin{equation}
P_{el,tot} = \sum_{1}^{4} AE_i
\end{equation}

The determination of how different consumers contribute to the total electric power demand is less trivial. The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc. None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.

\subsubsection*{HVAC systems}

For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/Pel_vs_time}
	\caption[Total electric power demand versus time (Daily average)]{Total electric power demand versus time (Daily average)}
	\label{fig:pelvstime}
\end{figure}

Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months. Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable. This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/PelRel_vs_time}
	\caption{Instantaneous power demand over daily average, winter versus summer day}
	\label{fig:pelrelvstime}
\end{figure}

Hence, the HVAC electric power demand was estimated as follows:
\begin{equation}
P_{el,HVAC} =
\begin{cases}
P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
0, & \text{otherwise}
\end{cases}
\end{equation}
where $P_{el,ref}(t)$ is calculated as:
\begin{equation}
	P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
\end{equation}

\subsubsection*{Thrusters}

When entering port areas, the ship needs to use thrusters to maneuver and berth. In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.

The observation of Figure XX shows how the power demand from thrusters can be identified as "spikes" in the total electric power demand. In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average. From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).

By comparing the reference profile to the instantaneous one, the points where the former is more than 10\% higher than the latter are identified as "thruster-on" points and treated consequently.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/Thruster_selection}
	\caption{Example of the procedure of selection of thruster power demand}
	\label{fig:thruster_selection}
\end{figure}
